---
title: "Statistical Inference Coursera Project - Part 2"
author: "Nick Horne"
date: "25 September 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Overview

Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package. This data looks at the `Effect of Vitamin C on Tooth Growth in Guinea Pigs`

#### 1. Load the ToothGrowth data and perform some basic exploratory data analyses

```{r}
# Load the data
data(ToothGrowth)

head(ToothGrowth)
summary(ToothGrowth)
```

#### 2. Provide a basic summary of the data.

The data contains 3 variables `len` - The length of odontoblasts (cells responsible for tooth growth). `supp` - Orange juice (OJ) or ascorbic acid (VC) or `dose` - The dose levels of vitamin C (mg/day).

```{r}
table(ToothGrowth$dose)

# Convert it to a factor
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```

The dose levels are either 0.5, 1 or 2 (mg/day) and are equaliy distrubuted amongs the 60 observations. There are 30 observations for each suppliment.

```{r}
aggregate(len~supp+dose, data=ToothGrowth, length)
```

There are 10 observations for each combination of dose and suppliment.

```{r}
plot <- ggplot(ToothGrowth, 
               aes(x=factor(dose),y=len,fill=factor(dose)))
plot + geom_boxplot(notch=F) + facet_grid(.~supp) +
     xlab("Dosage (mg per day)") +   
     ylab("Length of Tooth cell") +  
     ggtitle("Effect of Dosage and Supplement Type on Tooth Growth")
```

The graph shows that both suppliments increase tooth cell length with dosage with the orange juice being more effective at lower dosages. 

#### 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

```{r}
t.test(ToothGrowth[ToothGrowth$supp=="OJ",]$len, ToothGrowth[ToothGrowth$supp=="VC",]$len, paired = FALSE, var.equal = FALSE)
```

#### 4. State your conclusions and the assumptions needed for your conclusions.

Findings:

* Higher doses of vitamin C result in a longer total cell length regardless of the suppliment
* Orange juice is more effective at lower levels when compared to the ascorbic acid. 
* The orange juice effectiveness seems to tail off at higher dose levels

These findings are based on the assumptions:

* The guinea pigs were randomly assigned and representative of the entite population


